Concurrent Programming in Harmony

### 2021-03-01 book/paper.tex
Latest

The implementation in \autoref{fig:queueMS}~\cite{MS98} is similar but uses
separate locks for the head and tail, allowing a \texttt{put} and a \texttt{get} operation
to proceed concurrently.  To avoid contention between the head and the tail,
the queue uses a dummy node at the head of the linked list.
Except initially, the dummy node is the last node that was dequeued.
Note that neither the \texttt{head} nor \texttt{tail} pointer are
ever \texttt{None}.

>>>>

\autoref{fig:queueMS} shows another
concurrent queue implementation~\cite{MS96}.
It is well-known, but what is not often realized is that
it requires sequentially consistent memory, which is not
said explicitly in the paper.
As a result, the algorithm is not guaranteed to work
correctly with most modern programming languages and
computer hardware.  But it is still useful to study it.
The implementation uses separate
locks for the head and the tail,
allowing a \texttt{put} and a \texttt{get} operation
to proceed concurrently.  To avoid contention between the head and the tail,
the queue uses a dummy node at the head of the linked list.
Except initially, the dummy node is the last node that was dequeued.
Note that neither the \texttt{head} nor \texttt{tail} pointer are
ever \texttt{None}.
The problem is when the queue is empty and there are concurrent
\texttt{get} and \texttt{put} operations.
They obtain separate locks and then concurrently access the
\texttt{next} field in the dummy node---a data race with
undefined semantics in most environments.
